---
title: "A_2.final"
author: "Study group 3"
date: '2022-11-04'
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, error = TRUE)
```
#### Loading required packages

```{r}
pacman::p_load(glue,
       tidyr,
       data.table,
       moments,
       tidybayes,
       tibble,
       cowplot,
       viridis,
       brms,
       stringr,
       rstan,
       cmdstanr,
       magrittr,
       gridExtra,
       grid,
       lattice,
       ggplot2,
       ggridges,
       ellipse,
       Rmisc, 
       dplyr, 
       "rmarkdown", 
       knitr)
pacman::p_load(tidyverse)
pacman::p_load(purr)
pacman::p_load(MCMCglmm)
pacman::p_load(readxl)
pacman::p_load(metafor)
```

# Assignment 2: meta-analysis

## Questions to be answered

1.  Simulate data to setup the analysis and gain insight on the structure of the problem. Simulate one dataset of 100 studies (n of participants should follow a normal distribution with mean of 20, sd of 10, but no fewer than 10 participants), with a mean effect size of 0.4, average deviation by study of .4 and measurement error of .8. The data you get should have one row per study, with an effect size mean and standard error. Build a proper bayesian model to analyze the simulated data. Then simulate publication bias (only some of the studies you simulate are likely to be published, which?), the effect of publication bias on your estimates (re-run the model on published studies, assess the difference), and use at least one technique to assess publication bias. remember to use at least one plot to visualize your results. BONUS question: do a power/precision analysis.

2.  What is the current evidence for distinctive vocal patterns in schizophrenia? Use the data from Parola et al (2020) - <https://www.dropbox.com/s/0l9ur0gaabr80a8/Matrix_MetaAnalysis_Diagnosis_updated290719.xlsx?dl=0> - focusing on pitch variability (PITCH_F0SD). Describe the data available (studies, participants). Using the model from question 1 analyze the data, visualize and report the findings: population level effect size; how well studies reflect it; influential studies, publication bias. BONUS question: add the findings from <https://www.medrxiv.org/content/10.1101/2022.04.03.22273354v2>. BONUS question: assess the effect of task on the estimates (model comparison with baseline model)

# Sara
## Question 1

### Simulation

#### Outlining prior parameter provided by the assignment description
```{r}
mean_effect <- 0.4
effect_sd <- 0.4
meas_error <- 0.8
par_mean <- 20
par_sd <- 10
n <- 100
```

#### A simulation of participant data of multiple visits using the provided data
```{r}
set.seed(954)

sim_studies <-
  tibble(
    study_ID = seq(1:n),
    n_participants = 
        round(rtnorm(n, mean=par_mean, sd=par_sd, lower=10))
    ) 
      
  
for (i in seq(nrow(sim_studies))){
    sim_studies$study_effect[i] <- 
      rnorm(1,mean_effect,effect_sd)
    temp <- 
      rnorm(sim_studies$n_participants[i],sim_studies$study_effect[i], meas_error)
    sim_studies$mean_effect_size[i] <- 
      mean(temp)
    sim_studies$sd_effect[i] <- 
      sd(temp)
    sim_studies$standard_error[i] <-
      sim_studies$sd_effect[i]/sqrt(sim_studies$n_participants[i])
  }
```

### Bayesian model

#### A Bayesian model illustrating potential effect sizes on individual participants
```{r}
model_study <- bf(mean_effect_size|se(standard_error) ~1 + (1|study_ID))
```

### Priors

#### Generating prior data simulations to model, using parameters provided in class
```{r}
get_prior(data = sim_studies, family = gaussian, model_study)
priors <- c(
  prior(normal(0, 0.3), class=Intercept),
  prior(normal(0, 0.3), class=sd))
```

### Model

#### Modeling using sample_prior = 'only'
```{r}
model_prior <- brm(
  model_study, 
  data = sim_studies, 
  prior = priors,
  family = gaussian,
  refresh=0,
  sample_prior = 'only',
  iter=10000,
  warmup = 1000,
  backend = "cmdstanr",
  threads = threading(2),
  chains = 2,
  cores = 2,
  control = list(
    adapt_delta = 0.99,
    max_treedepth = 20
)
)
pp_check(model_prior, ndraws=100)
```
## Manuela

### Fitting model

#### Modeling the sampled priors along with the simulation
```{r}
model_prior_fit <- brm(
  model_study, 
  data = sim_studies, 
  prior = priors,
  family = gaussian,
  refresh=0,
  sample_prior = TRUE,
  iter=10000,
  warmup = 1000,
  backend = "cmdstanr",
  threads = threading(2),
  chains = 2,
  cores = 2,
  control = list(
    adapt_delta = 0.99,
    max_treedepth = 20
)
)
pp_check(model_prior_fit, ndraws=100)
```
#### Plotting and visualizing
```{r}
plot(model_prior_fit)
summary(model_prior_fit)
```



### Prior posterior update check

#### Plotting "prior-posterior update check on intercept" and "prior-posterior update check on standard deviation of the intercept"
```{r}
model_posterior <- as_draws_df(model_prior_fit)

plot1 <- ggplot(model_posterior)+
  geom_histogram(aes(prior_Intercept), fill='red', color='black', alpha=0.3, bins=50)+
  geom_histogram(aes(Intercept), fill='green', color='black', alpha=0.3, bins=50)+
  theme_classic()+
  ggtitle('prior-posterior update check on intercept')+
  xlab('intercept')
plot2 <- ggplot(model_posterior)+
  geom_histogram(aes(prior_sd_study_ID), fill='red', color='black', alpha=0.3, bins=50)+
  geom_histogram(aes(sd_study_ID__Intercept), fill='green', color='black', alpha=0.3, bins=50)+
  theme_classic()+
  ggtitle('prior-posterior update check on standard deviation of the intercept')+
  xlab('intercept')
grid.arrange(plot1, plot2)
```


### Simulation of publication bias, the effect of publication bias on our estimate and asses the publication bias (remember to visualize our results)

#### Simulating the effect size of the publication factors and filtering data for only published studies
```{r}
set.seed(843)

for (i in seq(nrow(sim_studies))){
  sim_studies$published[i] <-
    ifelse(abs(
      sim_studies$mean_effect_size[i])-(2*sim_studies$standard_error[i])>0 
      & sim_studies$mean_effect_size[i]>0,
      rbinom(1,1,0.9), rbinom(1,1,0.1))}

sim_studies <- sim_studies %>%
  mutate(published=as.factor(published))

pub_sim_studies <- dplyr::filter(sim_studies, published==1)


```

#### Modeling using sample_prior = 'only'
```{r}
pub_model_prior_fit <- brm(
  model_study, 
  data = pub_sim_studies, 
  prior = priors,
  family = gaussian,
  refresh=0,
  sample_prior = TRUE,
  iter=10000,
  warmup = 1000,
  backend = "cmdstanr",
  threads = threading(2),
  chains = 2,
  cores = 2,
  control = list(
    adapt_delta = 0.99,
    max_treedepth = 20
)
)
pp_check(pub_model_prior_fit, ndraws=100)
```

## Ditlev

#### Potting and assesing and transforming the brmsfit to draws
```{r}
plot(pub_model_prior_fit)
summary(pub_model_prior_fit)
pub_posterior <- as_draws_df(pub_model_prior_fit)
```

#### Plotting "prior-posterior update check on intercept" and "prior-posterior update check on standard deviation of the intercept"
```{r}
pub_plot1 <- ggplot(pub_posterior)+
  geom_histogram(aes(prior_Intercept), fill='red', color='black', alpha=0.3, bins=50)+
  geom_histogram(aes(Intercept), fill='green', color='black', alpha=0.3, bins=50)+
  theme_classic()+
  ggtitle('prior-posterior update check on intercept (published)')+
  xlab('intercept')
pub_plot2 <- ggplot(pub_posterior)+
  geom_histogram(aes(prior_sd_study_ID), fill='red', color='black', alpha=0.3, bins=50)+
  geom_histogram(aes(sd_study_ID__Intercept), fill='green', color='black', alpha=0.3, bins=50)+
  theme_classic()+
  ggtitle('prior-posterior update check on standard deviation of the intercept (published)')+
  xlab('sd')
grid.arrange(pub_plot1, pub_plot2)
```

#### Plotting "effect size with and without the un-published studis" and "tandard deviation of the effect size with and without the un-published studis"
```{r}
plot3 <- ggplot()+
  geom_histogram(aes(pub_posterior$Intercept), fill='red', color='black', alpha=0.3, bins=50)+
  geom_histogram(aes(model_posterior$Intercept), fill='green', color='black', alpha=0.3, bins=50)+
  theme_classic()+
  ggtitle('effect size with and without the un-published studis')+
  xlab('red=only published, green=with un-published')
plot4 <- ggplot()+
  geom_histogram(aes(pub_posterior$sd_study_ID__Intercept), fill='red', color='black', alpha=0.3, bins=50)+
  geom_histogram(aes(model_posterior$sd_study_ID__Intercept), fill='green', color='black', alpha=0.3, bins=50)+
  theme_classic()+
  ggtitle('standard deviation of the effect size with and without the un-published studis')+
  xlab('red=only published, green=with un-published')
grid.arrange(plot3, plot4)
```


## Question 2

#### Loading the data 
```{r}
matrix_ma <- read_excel("/Users/patrikmolnar/Desktop/Cognitive Science/Semester 3/Methods3/Matrix_MetaAnalysis.xlsx")
```

### Describing the data 

#### Filtering out NA & NR for SZ
```{r}
matrix_ma_filter_for_analysis <- matrix_ma %>%
  dplyr::filter(AGE_M_SZ!="NR") %>%
  dplyr::filter(AGE_M_SZ!="NA")
matrix_ma_filter_for_analysis <- matrix_ma_filter_for_analysis %>%
  dplyr::filter(AGE_SD_SZ!="NR") %>%
  dplyr::filter(AGE_SD_SZ!="NA")
matrix_ma_filter_for_analysis <- matrix_ma_filter_for_analysis %>%
  dplyr::filter(MALE_SZ!="NR") %>%
  dplyr::filter(MALE_SZ!="NA")
matrix_ma_filter_for_analysis <- matrix_ma_filter_for_analysis %>%
  dplyr::filter(FEMALE_SZ!="NR") %>%
  dplyr::filter(FEMALE_SZ!="NA")
```

#### Filtering out NA & NR for HC
```{r}
matrix_ma_filter_for_analysis <- matrix_ma_filter_for_analysis %>%
  dplyr::filter(AGE_M_HC!="NR") %>%
  dplyr::filter(AGE_M_HC!="NA")
matrix_ma_filter_for_analysis <- matrix_ma_filter_for_analysis %>%
  dplyr::filter(AGE_SD_HC!="NR") %>%
  dplyr::filter(AGE_SD_HC!="NA")
matrix_ma_filter_for_analysis <- matrix_ma_filter_for_analysis %>%
  dplyr::filter(MALE_HC!="NR") %>%
  dplyr::filter(MALE_HC!="NA")
matrix_ma_filter_for_analysis <- matrix_ma_filter_for_analysis %>%
  dplyr::filter(FEMALE_HC!="NR") %>%
  dplyr::filter(FEMALE_HC!="NA")
```

#### Making the variables numeric for SZ
```{r}
matrix_ma_filter_for_analysis$AGE_M_SZ <- as.numeric(matrix_ma_filter_for_analysis$AGE_M_SZ)
matrix_ma_filter_for_analysis$AGE_SD_SZ <- as.numeric(matrix_ma_filter_for_analysis$AGE_SD_SZ)
matrix_ma_filter_for_analysis$MALE_SZ <- as.numeric(matrix_ma_filter_for_analysis$MALE_SZ)
matrix_ma_filter_for_analysis$FEMALE_SZ <- as.numeric(matrix_ma_filter_for_analysis$FEMALE_SZ)
```

#### Making the variables numeric for HC
```{r}
matrix_ma_filter_for_analysis$AGE_M_HC <- as.numeric(matrix_ma_filter_for_analysis$AGE_M_HC)
matrix_ma_filter_for_analysis$AGE_SD_HC <- as.numeric(matrix_ma_filter_for_analysis$AGE_SD_HC)
matrix_ma_filter_for_analysis$MALE_HC <- as.numeric(matrix_ma_filter_for_analysis$MALE_HC)
matrix_ma_filter_for_analysis$FEMALE_HC <- as.numeric(matrix_ma_filter_for_analysis$FEMALE_HC)
```
## Patrik


#### Making both tibbles in order to combine them and make them easier for the eye
```{r}
a <- tibble(diagnosis = "SZ",
         mean_sample_size=mean(matrix_ma_filter_for_analysis$SAMPLE_SIZE_SZ),
         mean_numer_of_males=mean(matrix_ma_filter_for_analysis$MALE_SZ),
         mean_number_of_females=mean(matrix_ma_filter_for_analysis$FEMALE_SZ),
         mean_age=mean(matrix_ma_filter_for_analysis$AGE_M_SZ),
         mean_sd_age=mean(matrix_ma_filter_for_analysis$AGE_SD_SZ)
         )
  
b <- tibble(diagnosis = "HC",
         mean_sample_size=mean(matrix_ma_filter_for_analysis$SAMPLE_SIZE_HC),
         mean_numer_of_males=mean(matrix_ma_filter_for_analysis$MALE_HC),
         mean_number_of_females=mean(matrix_ma_filter_for_analysis$FEMALE_HC),
         mean_age=mean(matrix_ma_filter_for_analysis$AGE_M_HC),
         mean_sd_age=mean(matrix_ma_filter_for_analysis$AGE_SD_HC)
         )
```

#### Binding the rows together
```{r}
Demographic_overview <- bind_rows(a,b)
```

#### Showing the tibble
```{r}
Demographic_overview
```

#### Selceting the relevant variables
```{r}
matrix_pitch <- matrix_ma %>% 
  select('StudyID','Article','SAMPLE_SIZE_SZ','SAMPLE_SIZE_HC', 'PITCH_F0SD_HC_M','PITCH_F0SD_HC_SD','PITCH_F0SD_SZ_M','PITCH_F0SD_SZ_SD')
```

#### Filtering out the NA
```{r}
matrix_pitch <- matrix_pitch %>% 
  na.omit()
```

#### Merging diagnosis into one variable
```{r}
matrix_pitch <- matrix_pitch %>% 
  mutate(sample_size=(SAMPLE_SIZE_SZ+SAMPLE_SIZE_HC))
```

#### Creating IDs for the studies
```{r}
matrix_pitch <- matrix_pitch %>% 
  mutate(StudyID=as.factor(StudyID))
matrix_pitch <- matrix_pitch %>% 
  mutate(StudyID=as.numeric(StudyID))
matrix_pitch <- matrix_pitch %>% 
  mutate(StudyID=as.factor(StudyID))
```

#### Getting normalized results
```{r}
matrix_pitch <- escalc('SMD', 
n1i=SAMPLE_SIZE_HC,
n2i=SAMPLE_SIZE_SZ,
m1i = PITCH_F0SD_HC_M, 
m2i=PITCH_F0SD_SZ_M,
sd1i = PITCH_F0SD_HC_SD, 
sd2i = PITCH_F0SD_SZ_SD,
data = matrix_pitch)
matrix_pitch <- matrix_pitch %>% 
  rename(effect_size=yi)
```

#### Creating a loop to calculate sd effect size and se
```{r}
for (i in seq(nrow(matrix_pitch))){
  matrix_pitch$sd_effect[i] <- sqrt((sum((matrix_pitch$effect_size[i] - mean(matrix_pitch$effect_size))^2))/length(matrix_pitch))
  matrix_pitch$standard_error[i] <- matrix_pitch$sd_effect[i]/sqrt(matrix_pitch$sample_size)
}
```

#### Setting model
```{r}
model_matrix <- bf(effect_size|se(standard_error) ~1 + (1|StudyID))
```

#### Getting priors
```{r}
get_prior(data = matrix_pitch, family = gaussian, model_matrix)
```

#### Setting priors
```{r}
matrix_priors <- c(
  prior(normal( .3, 2.5), class=Intercept),
  prior(normal( 0, 2.5), class=sd))
```

#### Priors
```{r}
matrix_prior_fit <- brm(
  model_matrix, 
  data = matrix_pitch, 
  prior = matrix_priors,
  family = gaussian,
  refresh=0,
  sample_prior = 'only',
  iter=10000,
  warmup = 1000,
  backend = "cmdstanr",
  threads = threading(2),
  chains = 2,
  cores = 2,
  control = list(
    adapt_delta = 0.99,
    max_treedepth = 20
)
)
```

## Bryan

#### Assesing results
```{r}
matrix_prior_fit
```

#### PP check
```{r}
pp_check(matrix_prior_fit, ndraws=100)
```

#### Both data and priors

##### Including both data and priors
```{r}
matrix_fit <- brm(
  model_matrix, 
  data = matrix_pitch, 
  prior = matrix_priors,
  family = gaussian,
  refresh=0,
  sample_prior = 'only',
  iter=10000,
  warmup = 1000,
  backend = "cmdstanr",
  threads = threading(2),
  chains = 2,
  cores = 2,
  control = list(
    adapt_delta = 0.99,
    max_treedepth = 20
)
)
```

#### PP check
```{r}
pp_check(matrix_fit)
```

### Visualize and report 

#### Plotting and assesing the results
```{r}
plot(matrix_fit)
summary(matrix_fit)
```

#### Visualizing and assesing intercepts
```{r}
matrix_posterior <- as_draws_df(matrix_fit)
plot1 <- ggplot(matrix_posterior)+
  geom_histogram(aes(model_posterior$prior_Intercept), fill='red', color='black', alpha=0.3, bins=50)+
  geom_histogram(aes(Intercept), fill='green', color='black', alpha=0.3, bins=50)+
  theme_classic()+
  ggtitle('prior-posterior update check on intercept')+
  xlab('intercept')
```

#### Visualizing standard deviation
```{r}
plot2 <- ggplot(matrix_posterior)+
  geom_histogram(aes(model_posterior$prior_sd_study_ID), fill='red', color='black', alpha=0.3, bins=50)+
  geom_histogram(aes(model_posterior$sd_study_ID__Intercept), fill='green', color='black', alpha=0.3, bins=50)+
  theme_classic()+
  ggtitle('prior-posterior update check on standard deviation of the intercept')+
  xlab('intercept')
```

#### Printing plots
```{r}
plot1
plot2

grid.arrange(plot1, plot2)
```

### Influencial studies

#### Excluding the Cohen et al. (2014) by indexing
```{r}
excluded_matrix <- matrix_pitch %>% 
  dplyr::filter(StudyID!=6)
```

#### Running the model
```{r}
exclude_matrix_fit <- brm(
  model_matrix, 
  data = excluded_matrix, 
  prior = matrix_priors,
  family = gaussian,
  refresh=0,
  sample_prior = 'only',
  iter=10000,
  warmup = 1000,
  backend = "cmdstanr",
  threads = threading(2),
  chains = 2,
  cores = 2,
  control = list(
    adapt_delta = 0.99,
    max_treedepth = 20
)
)
```

#### Visualizing and plotting
```{r}
pp_check(exclude_matrix_fit)
plot(exclude_matrix_fit)
summary(exclude_matrix_fit)
```


